Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-15 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
many updates; summary table split

% ---------------------------------------------------
% Template. Complete when we have the main text done.
% ---------------------------------------------------
% This paper is organized as follows.
% %
% In Sec.~\ref{sec:review_theories} we present a short overview of the modified gravity theories
% which we will consider and (if any) the current constraints on them.
% %
% In Sec.~\ref{sec:method} we review the two frameworks we combine in our work: the {\sc Parspec}
% parametrization for the QNMs and the {\sc pSEOBNR} waveform model. We also explain how the
% theory-specific QNM calculation can be mapped onto the theory-agnostic {\sc Parspec} framework.
% %
% In Sec.~\ref{sec:results} we explain how we carry ...
% %
% Finally, we present our conclusion in Sec.~\ref{sec:discussion} and point directions for future work.
% %
% Unless stated otherwise, we use geometrical unit $c = G = 1$.

@@

A general procedure to describe deviations from the Kerr QNM frequencies $\omega_{J}^{\rm K}$ and
damping times $\tau_{J}^{\rm K}$ is to write,
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{J} &= \omega^{\rm K}_{J} \, (1 + \delta \omega_{J}), \\
\tau_{J}   &= \tau^{\rm K}_{J}   \, (1 + \delta \tau_{J}),
\end{align}
\end{subequations}
%
where $\delta\omega_{J}$ and $\delta\tau_{J}$ are the deformation parameters.
%
Here we use the subscript $J$ to represent the QNM labels $\{l,\, m,\, n\}$,
where $l$ and $m$ are the multipole indices, and $n$ the overtone number.
%
We will work exclusively with the $\ell = m = 2$ and $n = 0$ (fundamental mode)
QNM in this work, that is, $J = \{2,\, 2,\, 0\}$.
%
This approach was adopted, e.g., in Refs.~\cite{Gossan:2011ha,Meidam:2014jpa,Carullo:2018sfu}.

A caveat of the parametrizations of the form~\eqref{eq:general_deviation}, is that
$\delta\omega_{J}$ and $\delta\tau_{J}$ depend, in general, on the BH's source mass and spin.
%
Ideally, one would like to explicitly reinstate the dependence the source-frame
BH mass and spin for it would
%
%
(i) introduce deformation parameters which can be determined, once and for
all, from a specific gravity theory, and
%
(ii) make it easy to combine constraints from multiple source.

In the ParSpec framework this is accomplished by performing a bivariate
expansion of Eq.~\eqref{eq:general_deviation} in the detector-frame mass and
spin, namely
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{J} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{J} \, \left( 1 + \gamma \delta \omega^{(n)}_{J} \right), \\
\tau_{J}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{J}   \, \left( 1 + \gamma \delta \tau^{(n)}_{J} \right),
\end{align}
\end{subequations}
%
where $M$ and $\chi$ are the detector-frame mass and dimensionless spin of the
black hole; $\omega_{K}^{n}$, $\tau_{K}^{n}$ are dimensionless coefficients
of the spin expansion for Kerr QNMs; $\delta\omega_{K}^{n}$, $\delta\tau_{K}^{n}$ are
source-independent dimensionless coefficients that characterize the corrections to the Kerr QNM at each spin-order;
and finally $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on the beyond-GR modification is encapsulated in the dimensionless
parameter $\gamma$ given by,
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on the parameter $\ell$, with dimensions of length (and sets the
lengthscale at which beyond-GR modifications become important), and
the exponent $p$ which is related to how the beyond-GR modifications are
included to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made $\ell$ dimensionless by using the other
lenghtscale associated to the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}.

@@

@@

The polarization of the GW signal (in the observer's frame) are,
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical harmonics described by standard angular dependence indices $(l,m)$.\footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.} As our baseline model, we use the computationally efficient (time-domain) multipolar waveform model for quasicircular spin-aligned BBH mergers described in \cite{Mihaylov:2021bpf} (henceforth referred to as \SEOB{}\footnote{This waveform model is available in \texttt{LALSuite} \cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA} waveform approximant.}) which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$~\cite{Cotesta:2018fcv,Mihaylov:2021bpf}. The model uses an effective-one-body approach combined with a post-adiabatic solution of the equations of motion to describe the inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$. An accurate description of the merger is incorporated through calibration with NR simulations, as described in~\cite{Cotesta:2018fcv}, along with information for the merger and ringdown phases, from BH perturbation theory. The merger-ringdown waveform, $h_{l m}^\mathrm{merger-RD}$, is then stitched to inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) = h_{l m}^\mathrm{insp-plunge}\, \theta(t_\mathrm{match}^{l m} - t) + h_{l m}^\mathrm{merger-RD}\,\theta(t-t_\mathrm{match}^{l m})\,,
\end{equation}
%
where $\theta(t)$ is the Heaviside step function. The merger-ringdown waveform is expressed as an exponentially damped sinusoid ~\citep{Bohe:2016gbl,Cotesta:2018fcv,Mihaylov:2021bpf}
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})},
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} &= {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})\,,
\nonumber \\
&= \omega_{l m 0} - i / \tau_{l m 0}\,,
\end{align}
%
are the complex frequencies of the fundamental (n=0 overtone) QNMs of the remnant BH. The frequencies $\omega_{l m 0}$ and damping times $\tau_{l m 0}$ of the $(l, m)$ QNM can be read off from the real and imaginary parts of  $\sigma_{l m0}$ respetively. The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined in~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

In the $\SEOB$ model~\cite{Mihaylov:2021bpf}, the complex frequencies $\sigma_{l m 0}$ are computed by first determining the final mass and spin from estimates of the initial masses and spins through NR--fitting--formulas~\cite{Taracchini:2013rva,Hofmann:2016yih} and then converting them to the complex frequencies using BH perturbation theory--inspiral analytical fits outlined in~\cite{Berti:2005ys,Berti:2009kk}. Hence,

\begin{subequations}
\begin{eqnarray}
\omega_{l m 0}^{\text{GR}} &=& \omega_{l m 0}^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,\\
\tau _{l m 0}^{\text{GR}} &=& \tau _{l m 0}^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,.
\end{eqnarray}
\end{subequations}
where $(\omega_{l m 0}^{\text{GR}}, \tau _{l m 0}^{\text{GR}} )$ refer to the QNM predictions in baseline \SEOB{} model.  In this paper, however, we introduce a completely different expression of the QNM frequencies $(\omega_{l m 0}, \tau _{l m 0})$, as a series expansion on the spin of the remnant object $\chi_f$ [see Eqs.~\eqref{eq:kerr_expansion}].  For the rest of the paper, in keeping with the conventions introduced in Sec.~\ref{sec:review_parspec}, we are going to use the index $k$ to refer to $(l, m,n=0)$ mode. Accordingly, we are going to use $(\omega_k, \tau _k)$ to refer to $(\omega_{l m 0}, \tau _{l m 0})$ respectively.

In this formalism our QNM frequencies are expressed through functions,
%
\begin{subequations}
\begin{align}
\omega_k &= \omega_k(m_1, m_2, \chi_1, \chi_2,\ell, \{\delta \omega_k^{(j)}\})\\
\tau_k   &= \tau _k(m_1, m_2, \chi_1, \chi_2, \ell, \{\delta \tau_k^{(j)}\})
\end{align}
\end{subequations}
%
where the estimates of $(m_1,\, m_2,\, \chi_1,\, \chi_1)$ are used to predict the final mass and spin $(M_f,\, \chi_f)$ through~\cite{Taracchini:2013rva,Hofmann:2016yih}.
%
Additionally, $(\omega_k,\, \tau_k)$ depend on the coupling constant $\ell$
defined in Eq.~\eqref{eq:def_gamma} as well as the sets, $\{\delta \omega_k^{(j)}\},$ and
$\{\delta \tau_k^{(j)}\}$, where $j$ ranges from zero to the power of spin in Eqs.~\eqref{eq:kerr_expansion}
up to which we include corrections.
%
For example, if we restrict ourselves to corrections in the $j=0$, or $\chi_f^0$, terms for just the $k=(2,2,0)$ mode, then we would just have two additional degrees of freedom, $(\delta \omega_0^{(0)}, \delta \tau_0^{(0)})$. In this paper, we would restrict ourselves to corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$.

Using this parameterized waveform model, which we will call $\pSEOB$ in this paper, we infer bounds on our beyond-GR parameters for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}. We detail these results in Sec.~\ref{sec:results}.
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

Let us now establish the connection between theory-agnostic framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.
%
We begin by expanding $\omega$ and $\tau$, given by Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\label{eq:qnm_nongr_iso}
\begin{align}
M \omega_{220} &= \gamma \left( \delta\omega^{(0)} \omega^{(0)} + \chi \delta\omega^{(1)} \omega^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{j} \omega^{j}\,,
\\
\tau_{220}/M   &= \gamma \left( \delta\tau^{(0)} \tau^{(0)} + \chi \delta\tau^{(1)} \tau^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{j} \tau^{j}\,,
\end{align}
\end{subequations}
%
where, recall, we focus on the $\ell = m = 2$, $n=0$ QNM and have pulled out of
the sum all beyond-GR corrections, restricting ourselves to the nonspinning
($n=0$) and linear-order in spin ($n=1$) corrections to the Kerr QNMs.
%
These corrections are encoded in the coefficients $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$.

Just as in GR where the coefficients $\omega^{j}$, $\tau^{j}$ are determined by
comparison against the Kerr QNMs, a theory-specific calculation predicts the
numerical values of $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$
%
This means that using the $\pSEOB$ waveform model with theory-specific informed
QNMs we reduce the parameters space of beyond-GR parameters to one, i.e., to
only $\gamma$, which in turn contains information about the length scale $\ell$.
%
It is important to emphasize that our procedure is different from that of Ref.~\cite{Carullo:2021dui}
which, for a given value of $p$, varied all $\ell$, $\delta\omega^{(i)}$ and $\delta\tau^{(i)}$ parameters.

As we have seen in Sec.~\ref{sec:review_theories}, the QNM of slowly-rotating BHs
is modified gravity theories come in two families depending on their transformation upon a parity
transformation: axial and polar. Which is the one we use to match against Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work on a theory-by-theory case and in each theory perform a
a translation between the metric perturbations $h_{\mu\nu}$ in the Regger-Wheeler-Zerilli gauge and
connect it with the transverse-traceless gauge used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Sec. xxx),
%
In GR, both axial and polar QNMs are the same (a property known as \emph{isospectrality}) and therefore which
QNM we use to model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see the discussion
in Ref.~\cite{Hui:2021cpm} for a counterexample). and thus how axial and polar
gravitational QNMs appear in the GW signal has to be answered in
theory-by-theory.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode which, if excited during the merger, is
the most likely to appear in the signal.

% ---------------------------
% CONCRETE EXAMPLE OF MAPPING
% ---------------------------
Let us show one concrete example of how this mapping is done.
%
In Ref.~\cite{Wagle:2021tam} the QNMs of slow-rotating BHs in dCS gravity were calculated
and it was found that the axial gravitational modes have their damping time increased as we
increase the Chern-Simons coupling $\ell_{\rm CS}$ at fixed spin $\chi$ of the BH.
%
Having selected this branch of QNMs we can proceed to determine $\delta\omega^{(i)}$ and $\delta\tau^{(i)}$ as follows.
%
Using the fitting formula Eq.~(54a) of~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M \omega^{\rm CS}_{220} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi)^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part ${\rm Im}(\sigma_{220}) =  - 1 / \tau^{\rm CS}_{220}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm CS}^{4} / (M^4 \kappa)$ [thus $\kappa \zeta = (\ell_{\rm CS} / M)^4$]
and $c_{i}$ are fitting coefficients.
%
In the context of a binary BH merger, $M$ is interpreted as the remnant mass $\Mf$.

@@

>>>>

\begin{table}[th]
\begin{tabular}{c | c c}
\hline \hline
Theory & Constraint & This work \\
\hline
EdGB        & $\ell_{\rm GB} \leqslant 1.18$~km~(GW)~\cite{Lyu:2022gdr} & -- \\
dCS         & $\ell_{\rm CS} \leqslant 8.5$~km~(EM+GW)~\cite{Silva:2020acr}  & $\ell_{\rm CS} \leqslant 34.5$~km \\
cubic EFT   & \dots & \dots \\
quartic EFT & $\ell_{\rm qEFT} = \dots$~km~\cite{Sennett:2019bpc}  & \dots \\
\hline \hline
\end{tabular}
\caption{\hs{Mention it in the Introduction later, in a sort of `executive summary'.}}
\label{tab:bound_summary}
\end{table}

@@

A general procedure to describe deviations from the Kerr QNM frequencies $\omega_{k}^{\rm Kerr}$ and
damping times $\tau_{k}^{\rm Kerr}$ is to write,
%
\begin{subequations}
\label{eq:general_deviation}
\begin{align}
\omega_{k} &= \omega^{\rm Kerr}_{k} \, (1 + \delta \omega_{k}), \\
\tau_{k}   &= \tau^{\rm Kerr}_{k}   \, (1 + \delta \tau_{k}),
\end{align}
\end{subequations}
%
where $\delta\omega_{k}$ and $\delta\tau_{k}$ are the deformation parameters.
%
Here we use the subscript $k$ to represent the QNM labels $\{l,\, m,\, n\}$,
where $l$ and $m$ are the multipole indices, and $n$ the overtone number.
%
This approach was adopted, e.g., in Refs.~\cite{Gossan:2011ha,Meidam:2014jpa,Carullo:2018sfu}.

A caveat of parametrizations of the form~\eqref{eq:general_deviation}, is that
$\delta\omega_{k}$ and $\delta\tau_{k}$ depend, in general, on the BH's source mass and spin.
%
Ideally, one would like to explicitly reinstate the dependence the source-frame
BH mass and spin for it would
%
%
(i) introduce deformation parameters which can be determined, once and for
all, from a specific gravity theory, and
%
(ii) make it easy to combine constraints from multiple source.

In the ParSpec framework this is accomplished by performing a bivariate
expansion of Eq.~\eqref{eq:general_deviation} in the detector-frame mass and
spin, namely
%
\begin{subequations}
\label{eq:kerr_expansion}
\begin{align}
\omega_{k} &= \frac{1}{M} \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \omega^{(n)}_{k} \, \left( 1 + \gamma \delta \omega^{(n)}_{k} \right), \\
\tau_{k}   &= M     \sum_{n = 0}^{N_{\rm max}} \, \chi^{n} \tau^{(n)}_{k}   \, \left( 1 + \gamma \delta \tau^{(n)}_{k} \right),
\end{align}
\end{subequations}
%
where $M$ and $\chi$ are the detector-frame mass and dimensionless spin of the
black hole; $\omega_{k}^{n}$, $\tau_{k}^{n}$ are dimensionless coefficients
of the spin expansion for Kerr QNMs; $\delta\omega_{k}^{(n)}$, $\delta\tau_{k}^{(n)}$ are
source-independent dimensionless coefficients that characterize the corrections to the Kerr QNM at each spin-order;
and finally $N_{\rm max}$ is the order of the spin expansion.
%
All source dependence on the beyond-GR modification is encapsulated in the dimensionless
parameter $\gamma$ given by,
%
\begin{equation}
\gamma
%
= \left(\frac{\ell}{M_{\rm s}}\right)^{p}
%
= \left[
\frac{\ell c^2 (1 + z)}{G M}
\right]^{p},
\label{eq:def_gamma}
\end{equation}
%
which depends on the parameter $\ell$, with dimensions of length (and sets the
lengthscale at which beyond-GR modifications become important), and
the exponent $p$ which is related to how the beyond-GR modifications are
included to the Einstein-Hilbert action.
%
In Eq.~\eqref{eq:def_gamma} we made $\ell$ dimensionless by using the other
lenghtscale associated to the remnant black hole, i.e., its source-frame mass
$M_{\rm s}$, which we can also write in terms of the detector-frame mass $M$ through
the redshift $z$~\cite{Krolak:1987ofj}.

@@

We remark that in the GR limit, where $\delta\omega_{k}^{(n)} = \delta\tau_{k}^{(n)} = 0$ $\forall\, n$,
the series~\eqref{eq:kerr_expansion} truncated at $N_{\rm max} = 4$, reproduces with 1\% accuracy
the Kerr QNMs for spins $\chi \lesssim 0.7$.
%
The values of the fitting coefficients $\omega_{k}^{(n)}$ and $\tau_{k}^{(n)}$
can be found in Ref.~\cite{Maselli:2019mjd}, Table I.

@@

The polarization of the GW signal (in the observer's frame) are,
%
\begin{equation}
h_+(\iota,\varphi_0;t ) - i h_\times(\iota,\varphi_0;t) = \sum_{l, m} {}_{-\!2}Y_{l m}(\iota,\varphi_0)\, h_{l m}(t)\,,
\end{equation}
%
where ${}_{-\!2}Y_{l m}(\iota,\varphi_0)$ are the $-2$ spin-weighted spherical harmonics described by standard angular dependence indices $(l,m)$.\footnote{Note that we use $l$ to denote the angular dependence index and $\ell$ to denote the coupling constant in Sec.~\ref{sec:review_parspec}.} As our baseline model, we use the computationally efficient (time-domain) multipolar waveform model for quasicircular spin-aligned BBH mergers described in \cite{Mihaylov:2021bpf} (henceforth referred to as \SEOB{}\footnote{This waveform model is available in \texttt{LALSuite} \cite{lalsuite} as the \texttt{SEOBNRv4HM\_PA} waveform approximant.}) which contains the modes, $(l, |m|)=(2,2),(2,1)$, $(3,3)$, $(4,4)$, and $(5,5)$~\cite{Cotesta:2018fcv,Mihaylov:2021bpf}. The model uses an effective-one-body approach combined with a post-adiabatic solution of the equations of motion to describe the inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$. An accurate description of the merger is incorporated through calibration with NR simulations, as described in~\cite{Cotesta:2018fcv}, along with information for the merger and ringdown phases, from BH perturbation theory. The merger-ringdown waveform, $h_{l m}^\mathrm{merger-RD}$, is then stitched to inspiral-plunge waveform, $h_{l m}^\mathrm{insp-plunge}$ at a certain time $t = t^{\textrm{match}}_{l m}$, as
%
\begin{equation}
h_{l m}(t) = h_{l m}^\mathrm{insp-plunge}\, \theta(t_\mathrm{match}^{l m} - t) + h_{l m}^\mathrm{merger-RD}\,\theta(t-t_\mathrm{match}^{l m})\,,
\end{equation}
%
where $\theta(t)$ is the Heaviside step function. The merger-ringdown waveform is expressed as an exponentially damped sinusoid~\citep{Bohe:2016gbl,Cotesta:2018fcv,Mihaylov:2021bpf}
%
\begin{equation}
\label{RD}
h_{l m}^{\textrm{merger-RD}}(t) = \nu \ \tilde{A}_{l m}(t)\ e^{i \tilde{\phi}_{l m}(t)} \ e^{-i \sigma_{l m 0}(t-t^{\textrm{match}}_{l m})},
\end{equation}
%
where
%
\begin{align}
\sigma_{l m 0} &= {\rm Re}(\sigma_{l m 0}) + i \, {\rm Im}(\sigma_{l m 0})\,,
\nonumber \\
&= \omega_{l m 0} - i / \tau_{l m 0}\,,
\end{align}
%
are the complex frequencies of the fundamental ($n=0$ overtone) QNMs of the remnant BH.
%
The frequencies $\omega_{l m 0}$ and damping times $\tau_{l m 0}$ of the $(l,
m)$ QNM can be read off from the real and imaginary parts of $\sigma_{l m 0}$
respectively.
%
The functions $\tilde{A}_{l m}(t)$ and $\tilde{\phi}_{l m}(t)$ are defined in~\cite{Bohe:2016gbl,Cotesta:2018fcv}.

In the $\SEOB$ model~\cite{Mihaylov:2021bpf}, the complex frequencies $\sigma_{l m 0}$ are computed by first determining the final mass and spin from estimates of the initial masses and spins through NR--fitting--formulas~\cite{Taracchini:2013rva,Hofmann:2016yih} and then converting them to the complex frequencies using BH perturbation theory--inspiral analytical fits outlined in~\cite{Berti:2005ys,Berti:2009kk}. Hence,
%
\begin{subequations}
\begin{align}
\omega_{l m 0}^{\text{GR}} &= \omega_{l m 0}^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,,\\
\tau _{l m 0}^{\text{GR}} &= \tau _{l m 0}^{\text{GR}}(m_1, m_2, \chi_1, \chi_2)\,.
\end{align}
\end{subequations}
%
where $(\omega_{l m 0}^{\text{GR}}, \tau_{l m 0}^{\text{GR}} )$ refer to the
QNM predictions in baseline \SEOB{} model.
%
In this paper, however, we introduce a completely different expression of the
QNM frequencies $(\omega_{l m 0}, \tau_{l m 0})$, as a series expansion on the
spin of the remnant object $\chi_f$ [see Eqs.~\eqref{eq:kerr_expansion}].
%
For the rest of the paper, in keeping with the conventions introduced in
Sec.~\ref{sec:review_parspec}, we are going to use the index $k$ to refer to
$(l, m,n=0)$ mode. Accordingly, we are going to use $(\omega_k, \tau _k)$ to
refer to $(\omega_{l m 0}, \tau _{l m 0})$ respectively.

In this formalism our QNM frequencies are expressed through functions,
%
\begin{subequations}
\begin{align}
\omega_k &= \omega_k(m_1, m_2, \chi_1, \chi_2,\ell, \{\delta \omega_k^{(j)}\}),\\
\tau_k   &= \tau _k(m_1, m_2, \chi_1, \chi_2, \ell, \{\delta \tau_k^{(j)}\}),
\end{align}
\end{subequations}
%
where we fixed a certain value of $p$, and the estimates of $(m_1,\, m_2,\, \chi_1,\, \chi_1)$ are used to predict
the final mass and spin $(M_f,\, \chi_f)$ through~\cite{Taracchini:2013rva,Hofmann:2016yih}.
%
% HS: we implicitly fixed p here. Could we leave it free in PE study. This is similar
% to the 'number of dimensions'-like studies. Here it would tells us quite generically
% what is the dimension of the modification to the EH-action we include!
%
Additionally, $(\omega_k,\, \tau_k)$ depend on the coupling constant $\ell$
defined in Eq.~\eqref{eq:def_gamma} as well as the sets, $\{\delta
\omega_k^{(n)}\},$ and $\{\delta \tau_k^{(n)}\}$, where $n$ ranges from zero to
the power of spin in Eqs.~\eqref{eq:kerr_expansion} up to which we include
corrections.
%
For example, if we restrict ourselves to corrections with $n=0$, or
$\chi_f^0$, to the $k=(2,2,0)$ mode only, then we would just have two
additional parameters, $(\delta \omega_0^{(0)}, \delta \tau_0^{(0)})$.
%
In this paper, we will restrict ourselves to corrections to the $n=0$ and
$n=1$ terms, i.e., leading and next-to-leading order corrections to the spin,
$\chi_f$.

Using this parameterized waveform model, which we will call $\pSEOB$ in this
paper, we infer bounds on our beyond-GR parameters for specific cases of
modified gravity theories outlined in Sec.~\ref{sec:review_theories}. We detail
these results in Sec.~\ref{sec:results}.
%
\hs{Should we call this waveform differently to distinguish it from your previous papers?}
%
%In this paper, we will restrict ourselves to the $k=(2,2,0)$ QNM and corrections to the $j=0$ and $j=1$ terms, i.e., leading and next-to-leading order corrections to the spin, $\chi_f$. Assuming a prior which is uniform on the beyond-GR parameters, $(\ell, \{\delta \omega_k^{(j)}\},\{\delta \tau_k^{(j)}\})$, we employ a Bayesian framework to stochastically sample over the parameter space using the \texttt{LALInference} algorithm.~\cite{lallsuite}.
%
%The bounds on these beyond-GR parameters, for specific cases of modified gravity theories outlined in Sec.~\ref{sec:review_theories}, are detailed in Sec.~\ref{sec:results}.

@@

Let us now establish the connection between theory-agnostic framework of the
$\pSEOB$ waveform model and the theory-specific QNM calculations.
%
We begin by rewriting $\omega_{k}$ and $\tau_{k}$, given by Eqs.~\eqref{eq:kerr_expansion} as,
%
\begin{subequations}
\label{eq:qnm_nongr_iso}
\begin{align}
M \omega_{k} &= \gamma \left( \delta\omega^{(0)} \omega^{(0)} + \chi \delta\omega^{(1)} \omega^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{j} \omega^{j}\,,
\\
\tau_{k}/M   &= \gamma \left( \delta\tau^{(0)} \tau^{(0)} + \chi \delta\tau^{(1)} \tau^{(1)} \right)
+ \sum_{n=0}^{N_{\rm max}} \chi^{(n)} \tau^{(n)}\,,
\end{align}
\end{subequations}
%
We will work exclusively with the $\ell = m = 2$ and $n = 0$ (fundamental mode)
QNM, i.e., $k = \{2,\, 2,\, 0\}$. For this reason we will omit the subcripts
$k$ hereafter and no ambiguity in notation between the overtone index and the
subscript $n$ in the series should arise.
%
We have also pulled out from the sum all beyond-GR corrections, restricting
ourselves to the nonspinning ($n=0$) and linear-order in spin ($n=1$)
corrections to the Kerr QNMs.
%
% These corrections are encoded in the coefficients $\delta\omega^{(n)}$ and $\delta\tau^{(n)}$.

How can we determine the beyond-GR corrections? In GR, comparison between the numerically determined
Kerr QNMs against the fitting formula~\eqref{eq:qnm_nongr_iso} fixes the GR expansion coefficients
$\omega^{(n)}$, $\tau^{(n)}$.
%
We can proceed in a similar way with QNMs calculated in the context of a beyond-GR theory.
%
In particular, in the literature, we can already find fitting formulas relating
the QNMs to the BHs mass, spin and coupling constant $\ell$, the latter being
specific to each theory, up to $n=1$ in the spin expansion.
%
The idea is then to compare these formulas against Eq.~\eqref{eq:qnm_nongr_iso}
to fix $p$, $\delta\omega^{(n)}$, and $\delta\tau^{(n)}$.

This means that using the $\pSEOB$ waveform model with theory-specific informed
QNMs we reduce the parameter space of beyond-GR parameters to $\ell$ alone.
%
We emphasize that our procedure is different from that of
Ref.~\cite{Carullo:2021dui} which, for a given value of $p$, varied all $\ell$,
$\delta\omega^{(n)}$ and $\delta\tau^{(n)}$ parameters.

As we have seen in Sec.~\ref{sec:review_theories}, the QNM of slowly-rotating
BHs is modified gravity theories come in two families depending on their
transformation upon a parity transformation: axial and polar. Which is the one
we use to match against Eqs.~\eqref{eq:qnm_nongr_iso}?
%
To answer this question one has to work on a theory-by-theory case and in each
theory perform a translation between the metric perturbations $h_{\mu\nu}$ in
the Regger-Wheeler-Zerilli gauge and connect it with the transverse-traceless
gauge used to described GWs (see e.g.~Ref.~\cite{Maggiore:2018sht}, Sec. xxx),
%
In GR, both axial and polar QNMs are the same (a property known as \emph{isospectrality}) and therefore which
QNM we use to model the ringdown makes no difference.
%
In beyond-GR theories, isospectrality is in general broken (see the discussion
in Ref.~\cite{Hui:2021cpm} for a counterexample). and thus how axial and polar
gravitational QNMs appear in the GW signal has to be answered in
theory-by-theory.
%
This is outside the scope of this paper and here we take the more pragmatic
approach of simply choosing the \emph{least damped gravitational mode between
the two parities.}
%
The justification is this is the mode which, if excited during the merger, is
the most likely to appear in the signal.

We performed the mapping between theory-specific QNM calculation and the ParSpec framework
under the hypothesis above for the theories listed in Sec.~\ref{sec:review_theories}.
%
We leave the details of our calculation to Appendix~\ref{app:map_details} and
summarize our result in Table~\ref{tab:ref_theories_qnms}.

@@

\appendix

\section{Details of the determination of the theory-specific
ParSpec coefficients}
\label{app:map_details}

\subsection{dynamical Chern-Simons}
\label{app:map_dcs}

Let us show one concrete example of how this mapping is done.
%
In Ref.~\cite{Wagle:2021tam} the QNMs of slow-rotating BHs in dCS gravity were calculated
and it was found that the axial gravitational modes have their damping time increased as we
increase the Chern-Simons coupling $\ell_{\rm CS}$ at fixed spin $\chi$ of the BH.
%
Having selected this branch of QNMs we can proceed to determine $\delta\omega^{(i)}$ and $\delta\tau^{(i)}$ as follows.
%
Using the fitting formula Eq.~(54a) of~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M \omega^{\rm CS}_{220} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi)^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part ${\rm Im}(\sigma_{220}) =  - 1 / \tau^{\rm CS}_{220}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm CS}^{4} / (M^4 \kappa)$ [thus $\kappa \zeta = (\ell_{\rm CS} / M)^4$]
and $c_{i}$ are fitting coefficients.
%
In the context of a binary BH merger, $M$ is interpreted as the remnant mass $\Mf$.
